---
title: "Paris_green_space_processing"
author: "Jo Garrett"
date: '2024-02-22'
output: word_document
---

```{r libraries, echo = FALSE}
library(terra)
library(dplyr)
library(sf)
library(nngeo)

```

## 1. Read in data
## 2. Exclude spaces that are planned
```{r read_data, echo = FALSE}
setwd("E:\\Projects\\Users\\JoGarrett\\REGREEN_crib\\green-spaces-woods-open-or-planned-to-open-to-the-public")
greenspaces_raw<-vect("espaces-verts-et-boises-surfaciques-ouverts-ou-en-projets-douverture-au-public.shp") %>%
  makeValid()

## Exclude projects which are planned
ind<-which(greenspaces_raw$categlib!="Espace vert en projet")

greenspaces_open<-greenspaces_raw[ind,]
rm(ind, greenspaces_raw)
#writeVector(greenspaces_open, "greenspaces_open.shp")

```


## 3. Join together spaces that are <30 m apart
```{r within_30m, echo = FALSE}

## Add a 15 m buffer to each space
greenspaces_buff<-terra::buffer( greenspaces_open,width=15) %>%
  ## Combine overlapping spaces
  terra::aggregate() %>%
  ## fill any holes 
    fillHoles() %>%
  ## remove the 15 m buffer, such that spaces within 30 m are now joined 
  terra::buffer( width=(-15)) %>%
  ## disaggregate dataset such that there is an individual record for each polygon
  terra::disagg()


rm(greenspaces_open)
greenspaces_buff$area_terra<-expanse(greenspaces_buff)
```

## 4. Find the distance to the nearest green space
```{r find_nearest_distance, echo = FALSE}

greenspaces_sf <- sf::st_as_sf(greenspaces_buff)


nn <- st_nn(greenspaces_sf,y = greenspaces_sf, k = 2, maxdist = 500)  # Finding 2 nearest neighbors, the first one will be itself

# Custom function to extract second subelement or return NA
extract_second_element <- function(x) {
  if (length(x) == 2) {
    return(x[[2]])
  } else {
    return(NA)
  }
}

# Find index for nearest neighbour
nearest_neighbour <- sapply(nn, extract_second_element)
rm(nn,extract_second_element)

greenspaces_proj<-project(greenspaces_buff,"EPSG:2154")
# Calculate distance to nearest neighbour
for (i in 1:NROW(greenspaces_proj)){
    if (is.na(nearest_neighbour[i])==TRUE){
      greenspaces_proj$nearest_dist[i]<-501
    }else {
      greenspaces_proj$nearest_dist[i]<-terra::distance(greenspaces_proj[i,],greenspaces_proj[nearest_neighbour[i],])
    }
  }

rm(nearest_neighbour)

```

## 5. Find the distance to the nearest water body
```{r find_nearest_water_body, echo = FALSE}
water_bodies<-st_read("water_paris_polygon_osm.shp") %>%
  st_make_valid()

nn <- st_nn(greenspaces_sf,y = water_bodies, k = 1, maxdist = 500)  # Finding nearest neighbors

# Custom function to extract first subelement or return NA
extract_first_element <- function(x) {
  if (length(x) == 1) {
    return(x[[1]])
  } else {
    return(NA)
  }
}

## Find nearest neighbour
nearest_neighbour <- sapply(nn, extract_first_element )
rm(nn,extract_first_element )

water_vect<-vect(water_bodies) %>%
  project("epsg:2154")
  

## Find distance to nearest neighbour
greenspaces_proj$water_dist<-NA
for (i in 1:NROW(greenspaces_proj)){
    if (is.na(nearest_neighbour[i])==TRUE){
      greenspaces_proj$water_dist[i]<-501
    }else {
      greenspaces_proj$water_dist[i]<-terra::distance(greenspaces_proj[i,],water_vect[nearest_neighbour[i],])
    }
  }

rm(nearest_neighbour)

```

## 6. Select green spaces within the four central departments of Paris
```{r read_central, echo =FALSE}

setwd("E:\\Projects\\Users\\JoGarrett\\REGREEN_crib\\Analysis_2024")
central_region<-vect("central_departments.shp") 
crs(central_region)<-"EPSG:2154"

## Dissolve the four central departments into one polygon
central_dis<-terra::aggregate(central_region)

## Find where green spaces intersect or are within the four central departments
relates_ind<-relate(greenspaces_proj, central_dis, relation=c("intersects", "within"))
greenspaces_central<-greenspaces_proj[relates_ind] %>%
  makeValid()
rm(relates_ind)

rm(central_dis, central_region)

writeVector(greenspaces_central, "greenspaces_central.shp")
```





