---
title: "step_3_upscaling"
author: "Jo Garrett"
date: '2024-04-23'
output: word_document
---
```{r libraries,echo = FALSE}
## Make sure all libraries are installed and set working directory paths to own locations
library(terra)
library(dplyr)
library(ggplot2)
library(cowplot)
library(flextable)
library(zoo)
library(readr)
library(tidyterra)
library(RColorBrewer)

data_d<-"E:\\Projects\\Users\\JoGarrett\\REGREEN_crib\\Analysis_2024"
figure_d<-"E:\\Projects\\Users\\JoGarrett\\REGREEN_crib\\Analysis_2024\\Figs_for_paper"
```

## Read in data

```{r read_parks, echo = FALSE}

## Set working directory
setwd(data_d)

## Load in parks data using terra package
Parks_grtr_1000<-vect("Parks_grtr_1000_with_preds.shp")

parks_bounding<-ext(Parks_grtr_1000) %>%
  as.polygons() %>%
  buffer(width = 1000)
```



```{r read_temp_data, echo =FALSE}


## Set working directory
setwd(data_d)
## Read in data
temp_proj_191<-rast("Temp_data\\2019191_TAVG.tif") %>%
  ## project to the same coordinate reference system as the green spaces
  terra::project( "EPSG:2154") %>%
## Crop to the bounding box 
  crop( parks_bounding) %>%
## Resample by a factor of 6 (to every 5 m)
  disagg( fact= 6)

temp_proj_203<-rast("Temp_data\\2019203_TAVG.tif") %>%
  terra::project( "EPSG:2154") %>%
  crop( parks_bounding) %>%
  disagg( fact= 6)

temp_proj_205<-rast("Temp_data\\2019205_TAVG.tif") %>%
  terra::project( "EPSG:2154") %>%
  crop( parks_bounding) %>%
  disagg( fact= 6)

rm(parks_bounding)
```

## Calculate the boundary temperatures for all green spaces in Paris for the three hot days

```{r boundary_temp_cals, echo = FALSE}

# Create empty data frame
boundary_temp<-data.frame(boundary_temp_191= numeric(NROW(Parks_grtr_1000)), boundary_temp_203= numeric(NROW(Parks_grtr_1000)), boundary_temp_205= numeric(NROW(Parks_grtr_1000)))

# Run processing for all green spaces at least 1,000m2
for (i in 1:NROW(Parks_grtr_1000)){
  # Calculate greenspace buffer 15 m inside the boundary
    Buff_min<-terra::buffer(Parks_grtr_1000[i,], width=-15)
    # Where the area of the above is 0, try with 10 m inside boundary
    if (expanse(Buff_min) == 0){
       Buff_min<-terra::buffer(Parks_grtr_1000[i,], width=-10)
    }
    # Where the area of the above is 0, try with 5 m inside boundary
     if (expanse(Buff_min) == 0){
       Buff_min<-terra::buffer(Parks_grtr_1000[i,], width=-5)
     }
    # calculate buffer ring
    Buff_ring<-terra::symdif(Parks_grtr_1000[i,],Buff_min)
    # Include buffer rings only where they have value of 1m2 or more
   if(NROW(Buff_ring)>1){
      Buff_area<-expanse(Buff_ring)
      Buff_ring<-Buff_ring[Buff_area>1]
      rm(Buff_area)
   }
    rm(Buff_min)
    # Calculate mean temperature withint he buffer for each of the three hot days
      boundary_temp$boundary_temp_191[i]<-as.numeric(terra::extract(temp_proj_191, Buff_ring, fun = mean)[2])
    boundary_temp$boundary_temp_203[i]<-as.numeric(terra::extract(temp_proj_203, Buff_ring, fun = mean)[2])
    boundary_temp$boundary_temp_205[i]<-as.numeric(terra::extract(temp_proj_205, Buff_ring, fun = mean)[2])
    rm(Buff_ring)
}

Parks_grtr_1000$bt_191<-boundary_temp$boundary_temp_191

Parks_grtr_1000$bt_203<-boundary_temp$boundary_temp_203
Parks_grtr_1000$bt_205<-boundary_temp$boundary_temp_205

rm(boundary_temp)
```

### Plot the boundary temperatures

```{r plot_temps_by_park, echo = FALSE}

# Prepare plot for each of the three hot days
brks<-seq(16, 29, 0.5)
h<-hist(Parks_grtr_1000$bt_191, right = FALSE, breaks = brks)
h_df<-data.frame("Mids"=h$mids,"Counts"=h$counts) %>%
  mutate("Percent"=Counts/sum(Counts)*100)
p1<-ggplot(data = h_df, aes(x = Mids, y = Percent))+
  geom_col()+
  theme_minimal() +
  xlab("Greenspace boundary temperature (°C)")+
  ylab("%age of green spaces")+
  ggtitle("10th July")+
  scale_x_continuous(breaks = seq(16, 28.5, 1), limits =c(min(brks), max(brks)+0.5)) +
  ylim(c(0, 25))

h<-hist(Parks_grtr_1000$bt_203, right = FALSE, breaks = brks)
h_df<-data.frame("Mids"=h$mids,"Counts"=h$counts) %>%
  mutate("Percent"=Counts/sum(Counts)*100)
p2<-ggplot(data = h_df, aes(x = Mids, y = Percent))+
  geom_col()+
  theme_minimal() +
  xlab("Greenspace boundary temperature (°C)")+
  ylab("%age of green spaces")+
  ggtitle("22nd July")+
  scale_x_continuous(breaks = seq(16, 28.5, 1), limits =c(min(brks), max(brks)+0.5))+
  ylim(c(0, 25))

h<-hist(Parks_grtr_1000$bt_205, right = FALSE, breaks = brks)
h_df<-data.frame("Mids"=h$mids,"Counts"=h$counts) %>%
  mutate("Percent"=Counts/sum(Counts)*100)
p3<-ggplot(data = h_df, aes(x = Mids, y = Percent))+
  geom_col()+
  theme_minimal() +
  xlab("Greenspace boundary temperature (°C)")+
  ylab("%age of green spaces")+
  ggtitle("24th July")+
    scale_x_continuous(breaks = seq(16, 28.5, 1), limits =c(min(brks), max(brks)+0.5))+
  ylim(c(0, 25))

plot_grid(p1, p2, p3, nrow = 2, ncol = 2, align = "hv")
setwd(figure_d)
ggsave("temps_boundaries_three_days.png", dpi = 300, width = 7, height = 4)

rm(p1, p2, p3, h, h_df, brks)

## Present boundary temperature descriptive statistics
boundary_temp_191<-data.frame("Temp"=Parks_grtr_1000$bt_191, 
                              "Day" = 191)
boundary_temp_203<-data.frame("Temp"=Parks_grtr_1000$bt_203, 
                              "Day" = 203)
boundary_temp_205<-data.frame("Temp"=Parks_grtr_1000$bt_205, 
                              "Day" = 205)
boundary_temps_3_days<-rbind(boundary_temp_191, boundary_temp_203, boundary_temp_205)
rm(boundary_temp_191, boundary_temp_203, boundary_temp_205)

boundary_stats<-boundary_temps_3_days %>%
  group_by(Day) %>%
  summarise('Mean' = mean(Temp), 'Median' = median(Temp),'Min'=min(Temp), 'Max'= max(Temp)) %>% 
  flextable() %>%
  colformat_double( digits = 2)

  boundary_stats
  rm(boundary_stats)
```


### Read in population data

```{r population_data, echo = FALSE}
## Population data

setwd(data_d)

Pop<-vect("densibati_central_ile_de_france.shp") %>%
  makeValid()


```

### Read in relative risk of mortality 
```{r read_Rr, echo =FALSE}
## Read in the relative risks predicting mortality by 1 degree C
setwd(data_d)
RRs_by_temp<-read.csv("RR_interped_18102021.csv")
RRs_by_temp$X<-NULL

# Plot relative risks
p<-ggplot(RRs_by_temp, aes(x = Temperature, y = RR))+
  geom_point(size = 0.9)+
  theme_minimal()+
  ylim(c(0.99, max(pretty(RRs_by_temp$RR))))+
  geom_vline(xintercept = 21.6, linetype = "dashed", color = "red") +
  ylab("Relative risk of mortality")+
  xlab("Temperature (°C)")

setwd(figure_d)
ggsave("RR_mortality_by_temp.png", dpi = 300, height = 3, width = 3.5)
rm(p)

## Mortality rate Paris
mortality_rate_2019_paris<-6.065503/365
# Value of a statistical life

VSL = 4680000
```

## Calculate lives saved and associated economic value for the three hot days
```{r, calc_lives_saved,echo =FALSE}

Parks_grtr_1000$ls_191<-NA
Parks_grtr_1000$ls_203<-NA
Parks_grtr_1000$ls_205<-NA

ind<-which(names(Parks_grtr_1000)=="pred_cooli")
names(Parks_grtr_1000)[ind]<-"pred_cooling"
rm(ind)

Pop$area_calc<-terra::expanse(Pop)

# truncate cooling at 0
ind<-which(Parks_grtr_1000$pred_cooling<0)
Parks_grtr_1000$pred_cooling[ind]<-0
rm(ind)

# 
for (i in 1:NROW(Parks_grtr_1000)){
  if (Parks_grtr_1000$pred_cooling[i]>=15){

  # create a sequence of values to the maximum cooling distance
  Buff_vals<-seq(0, Parks_grtr_1000$pred_cooling[i], 30)
  # Add an extra interval if the maximum cooling distance is more than 15 m greater than maximum interval value
  if (Parks_grtr_1000$pred_cooling[i]>(max(Buff_vals)+15)){
    Buff_vals<-c(Buff_vals, max(Buff_vals)+30)
  }
    
  ## convert to two sequences for minimum and maximum value
  Buff_min_vals<-Buff_vals[1:(length(Buff_vals)-1)]
  Buff_max_vals<-Buff_vals[2:length(Buff_vals)]
    # find the temperature at the maximum cooling distance for day 191
  temp_at_max_dist_191<-Parks_grtr_1000$bt_191[i]+Parks_grtr_1000$pred_temp_[i]
  # calculate a sequence of temperature values from the boundary temperature to the temp at the cooling distance at the locations of the intervals (0, 30 m etc)
  temps_191<-seq(  Parks_grtr_1000$bt_191[i],   temp_at_max_dist_191, length.out = length(Buff_vals))
  # Calculate the mean values of the temperatures at the intervals such that they'e assocaited with the distances 15 m, 45 m etc
  temps_191<-rollapply(temps_191, 2, mean, by = 1, align = "left", partial = TRUE, fill = NA)[1:length(Buff_min_vals)]
  temp_at_max_dist_203<-Parks_grtr_1000$bt_203[i]+Parks_grtr_1000$pred_temp_[i]
  temps_203<-seq(  Parks_grtr_1000$bt_203[i],   temp_at_max_dist_203, length.out = length(Buff_vals))
  temps_203<-rollapply(temps_203, 2, mean, by = 1, align = "left", partial = TRUE, fill = NA)[1:length(Buff_min_vals)]
  temp_at_max_dist_205<-Parks_grtr_1000$bt_205[i]+Parks_grtr_1000$pred_temp_[i]
  temps_205<-seq(  Parks_grtr_1000$bt_205[i],   temp_at_max_dist_205, length.out = length(Buff_vals))
  temps_205<-rollapply(temps_205, 2, mean, by = 1, align = "left", partial = TRUE, fill = NA)[1:length(Buff_min_vals)]
  rm(Buff_vals)
  # Create empty variables
  lives_saved_191<-numeric(length(Buff_min_vals))
  lives_saved_203<-numeric(length(Buff_min_vals))
  lives_saved_205<-numeric(length(Buff_min_vals))
  
  for (j in 1:length(Buff_min_vals)){
    # Calculate the number of people living within each buffer
    Buff_min<-terra::buffer(Parks_grtr_1000[i,], width=Buff_min_vals[j])
    Buff_max<-terra::buffer(Parks_grtr_1000[i,], width=Buff_max_vals[j])
  Buff_ring<-terra::symdif(Buff_max,Buff_min)
    rm(Buff_max, Buff_min)
    Pop_inter<-terra::intersect(Pop,Buff_ring)
    Pop_inter$NewArea<-terra::expanse(Pop_inter)
    Pop_inter$New_Pop<-as.numeric(Pop_inter$POPMEN2016*(Pop_inter$NewArea/Pop_inter$area_calc))
      Buff_pop<-sum(Pop_inter$New_Pop)
      # Find the relative risk associated with the temperature they're exposed to for each of the three hot days
      ind_191<-which.min(abs(RRs_by_temp$Temperature - temps_191[j]))
      # calculate the expected deaths 
      deaths_at_temp_191<- ((mortality_rate_2019_paris*RRs_by_temp$RR[  ind_191])/1000)*Buff_pop
      ind_203<-which.min(abs(RRs_by_temp$Temperature - temps_203[j]))
      deaths_at_temp_203<- ((mortality_rate_2019_paris*RRs_by_temp$RR[  ind_203])/1000)*Buff_pop
      ind_205<-which.min(abs(RRs_by_temp$Temperature - temps_205[j]))
      deaths_at_temp_205<- ((mortality_rate_2019_paris*RRs_by_temp$RR[  ind_205])/1000)*Buff_pop
      rm(ind_191,ind_203, ind_205)
      # find the relative risk of mortality at the peak temperatures (therefore assuming no cooling)
      ind_pk_191<-which.min(abs(RRs_by_temp$Temperature - (Parks_grtr_1000$bt_191[i]+Parks_grtr_1000$pred_temp_[i])))
      ind_pk_203<-which.min(abs(RRs_by_temp$Temperature - (Parks_grtr_1000$bt_203[i]+Parks_grtr_1000$pred_temp_[i])))
  
      ind_pk_205<-which.min(abs(RRs_by_temp$Temperature - (Parks_grtr_1000$bt_205[i]+Parks_grtr_1000$pred_temp_[i])))
      # Calculate the expected deaths at thepeak temperatures for each of the three hot days
      deaths_at_pk_temp_191<- ((mortality_rate_2019_paris*RRs_by_temp$RR[ind_pk_191])/1000)*Buff_pop
      deaths_at_pk_temp_203<- ((mortality_rate_2019_paris*RRs_by_temp$RR[ind_pk_203])/1000)*Buff_pop
      deaths_at_pk_temp_205<- ((mortality_rate_2019_paris*RRs_by_temp$RR[ind_pk_205])/1000)*Buff_pop
      # Calculate the deaths prevented by subtracting the deaths at cooled temperatures from the expected deaths at peak temperature
      lives_saved_191[j]<-deaths_at_pk_temp_191- deaths_at_temp_191
      lives_saved_203[j]<-deaths_at_pk_temp_203- deaths_at_temp_203
      lives_saved_205[j]<-deaths_at_pk_temp_205- deaths_at_temp_205
    print(j)
  }
  # Calculate the total of lives saved for each greenspace for each of the three hot days
    Parks_grtr_1000$ls_191[i]<-sum(lives_saved_191)
    Parks_grtr_1000$ls_203[i]<-sum(lives_saved_203)
    Parks_grtr_1000$ls_205[i]<-sum(lives_saved_205)

    rm(lives_saved_191, lives_saved_203, lives_saved_205)
    print(i)
  } else{
    Parks_grtr_1000$ls_191[i]<-0
    Parks_grtr_1000$ls_203[i]<-0
    Parks_grtr_1000$ls_205[i]<-0
  }
}

# Calculate the value by multiplying the number of lives saved by the VSL
Parks_grtr_1000$v_191<-Parks_grtr_1000$ls_191*VSL
Parks_grtr_1000$v_203<-Parks_grtr_1000$ls_203*VSL
Parks_grtr_1000$v_205<-Parks_grtr_1000$ls_205*VSL

rm(i, j, deaths_at_pk_temp_191, deaths_at_pk_temp_203, deaths_at_pk_temp_205, deaths_at_temp_191, deaths_at_temp_203, deaths_at_temp_205, ind_pk_191, ind_pk_203, ind_pk_205, Pop_inter,  temp_at_max_dist_191, temp_at_max_dist_203, temp_at_max_dist_205, temps_191, temps_203, temps_205, Buff_max_vals, Buff_min_vals, Buff_pop, Buff_ring)

```


## Finding all days above 21.6
```{r read_orly_temps, echo = FALSE}

setwd(data_d)
# Read in temperatures for 2019 as recorded at Paris Orly airport
orly_temp<-read.csv("Paris_Orly_temp.csv")
# find where the average temperature was above the minimum mortality threshold
ind<-which(orly_temp$TAVG>21.6)
dates_above_mmt<-orly_temp[ind,]
rm(ind)
 dates_above_mmt$dates_as_date<- as.Date(dates_above_mmt$DATE,format = "%d/%m/%Y")

ind<-which(dates_above_mmt$Day_no==191 | dates_above_mmt$Day_no==203 | dates_above_mmt$Day_no==205)

## find the nearest day in temperatre compare to the three hot days already analysed
dates_above_mmt$diff_191<-numeric(NROW(dates_above_mmt))
dates_above_mmt$diff_203<-numeric(NROW(dates_above_mmt))
dates_above_mmt$diff_205<-numeric(NROW(dates_above_mmt))
for (i in 1:NROW(dates_above_mmt)){
  dates_above_mmt$diff_191[i]<-dates_above_mmt$TAVG[i]-dates_above_mmt$TAVG[ind[1]]
    dates_above_mmt$diff_203[i]<-dates_above_mmt$TAVG[i]-dates_above_mmt$TAVG[ind[2]]
    dates_above_mmt$diff_205[i]<-dates_above_mmt$TAVG[i]-dates_above_mmt$TAVG[ind[3]]
}

dates_above_mmt$min_diff<-character(NROW(dates_above_mmt))
for (i in 1:NROW(dates_above_mmt)){
  ind_min<-which.min(abs(dates_above_mmt[i,c(12:14)]))
  dates_above_mmt$min_diff[i]<-names(ind_min)
  rm(ind_min)
}

ind2<-which(dates_above_mmt$Day_no!=191 & dates_above_mmt$Day_no!=203 & dates_above_mmt$Day_no!=205  )

dates_above_mmt<-dates_above_mmt[ind2,]
rm(ind2)
```

### Find the boundary temperatures for all green spaces for all 40 hot days

```{r valuing_all_days, echo = FALSE}

###calculate boundary temperatures for all other 37 days above the MMT
for (i in 1:NROW(dates_above_mmt)) {
  column_name <- paste0("Temp_Day_", dates_above_mmt$Day_no[i])
  Parks_grtr_1000[[column_name]] <- numeric(NROW(Parks_grtr_1000))
}

rm(column_name)

# Finds the boundary temperature by adding the difference in average temperature between each day and the closest of the original three hot days to the boundary to the boundary temperature calculated for the closest of the original three hot days to the boundary 
column_names<-names(Parks_grtr_1000)[(NCOL(Parks_grtr_1000)-36):NCOL(Parks_grtr_1000)]
for (i in 1:NROW(dates_above_mmt)){
  day_col<-paste0("diff_",parse_number(dates_above_mmt$min_diff[i]))
  day_col_parks<-paste0("bt_",parse_number(dates_above_mmt$min_diff[i]))
Parks_grtr_1000[[column_names[i]]]<-Parks_grtr_1000[[day_col_parks]]+(dates_above_mmt[[day_col]][i])
  }


```



```{r calculate_lives_saved_value_all_days, echo = FALSE}
## Creating custom functions

# Create empty columns
for (i in 1:NROW(dates_above_mmt)) {
  column_name <- paste0("lives_saved_", dates_above_mmt$Day_no[i])
  Parks_grtr_1000[[column_name]] <- numeric(NROW(Parks_grtr_1000))
}

for (i in 1:NROW(dates_above_mmt)) {
  column_name <- paste0("value_", dates_above_mmt$Day_no[i])
  Parks_grtr_1000[[column_name]] <- numeric(NROW(Parks_grtr_1000))
}

# where x is the  Parks_grtr_1000$pred_cooling[i]
create_buff_dists<-function(x){
   # create a sequence of values to the maximum cooling distance
  Buff_vals<-seq(0,x, 30)
  # Add an extra interval if the maximum cooling distance is more than 15 m greater than maximum interval value
  if (x>(max(Buff_vals)+15)){
    Buff_vals<-c(Buff_vals, max(Buff_vals)+30)
  }
  return(Buff_vals)
}
Pop$area_calc<-expanse(Pop)


## Calculate populatiions within each 30 m buffer from the boundary to the maximum cooling distance
calc_buff_pops<-function(Parks_shapefile, Buff_vals, Pop){
  Buff_min_vals<-Buff_vals[1:(length(Buff_vals)-1)]
  Buff_max_vals<-Buff_vals[2:length(Buff_vals)]
  Buff_pop<-numeric(length(Buff_max_vals))
  for (d in 1:length(Buff_min_vals)){
    Buff_min<-terra::buffer(Parks_shapefile, width=Buff_min_vals[d])
    Buff_max<-terra::buffer(Parks_shapefile, width=Buff_max_vals[d])
    Buff_ring<-terra::symdif(Buff_max,Buff_min)
    rm(Buff_max, Buff_min)
    Pop_inter<-terra::intersect(Pop,Buff_ring)

    Pop_inter$NewArea<-expanse(Pop_inter)
    Pop_inter$New_Pop<-as.numeric(Pop_inter$POPMEN2016*(Pop_inter$NewArea/Pop_inter$area_calc))
    Buff_pop[d]<-sum(Pop_inter$New_Pop)
    rm(Buff_ring, Pop_inter)
  }
  return(Buff_pop)
}

#Calculate deaths prevented
for (i in 1:NROW(Parks_grtr_1000)){
  
  Buff_vals<-create_buff_dists( Parks_grtr_1000$pred_cooling[i])
  if ( Parks_grtr_1000$pred_cooling[i]>=15){
    Buff_pop<-calc_buff_pops(Parks_grtr_1000[i,], Buff_vals, Pop)
    Buff_min_vals<-Buff_vals[1:(length(Buff_vals)-1)]
    Buff_max_vals<-Buff_vals[2:length(Buff_vals)]
    for (j in 1:NROW(dates_above_mmt)){
      # find the temperature at the maximum cooling distance
      col_ind<-which(match(names(Parks_grtr_1000), column_names[j])==1)
      temp_at_boundary<-values(Parks_grtr_1000[i,col_ind]) %>%
        unlist()
      temp_at_max_dist<-temp_at_boundary+Parks_grtr_1000$pred_temp_[i]
      temps_seq<-seq( temp_at_boundary,   temp_at_max_dist, length.out = length(Buff_vals))
      temps_seq_mean<-rollapply(temps_seq, 2, mean, by = 1, align = "left", partial = TRUE, fill = NA)[1:length(Buff_min_vals)]
      rm(temps_seq)
      abs_diffs <- sapply(temps_seq_mean, function(x) abs(RRs_by_temp$Temperature - x))
      ind_temp <- apply(abs_diffs, 2, which.min)
      # Calculate expected deaths at each temperature within each 30 m buffer
      deaths_at_temp<- ((mortality_rate_2019_paris*RRs_by_temp$RR[  ind_temp])/1000)*Buff_pop
      rm(ind_temp)
      # Calculate expected deaths at the peak temperature (as  if there was no cooling)
      ind_pk<-which.min(abs(RRs_by_temp$Temperature - temp_at_max_dist))
      deaths_at_pk_temp<- ((mortality_rate_2019_paris*RRs_by_temp$RR[ind_pk])/1000)*Buff_pop
      # Calculate deaths prevented for each 30 m buffer 
      lives_saved_seq<-deaths_at_pk_temp- deaths_at_temp
      # Calculate value by multiplying the deaths prevented by the VSL
      value_seq<-lives_saved_seq*VSL
      column_name_lives<-gsub("Temp_Day_", "lives_saved_",column_names[j])
      col_ind<-which(names(Parks_grtr_1000)==column_name_lives)
      # Calculate the total deaths prevented and lives saved for each green space
      Parks_grtr_1000[i, col_ind]<-sum(lives_saved_seq, na.rm =TRUE)
      column_name_value<-gsub("Temp_Day_", "value_",column_names[j])
      col_ind<-which(names(Parks_grtr_1000)== column_name_value)
      Parks_grtr_1000[i, col_ind]<-sum(value_seq, na.rm = TRUE)
      
      rm(lives_saved_seq, value_seq, column_name_lives, column_name_value, ind_pk, deaths_at_pk_temp, deaths_at_temp, abs_diffs, temp_at_boundary, temp_at_max_dist, temps_seq_mean)
      print(paste("j", j))
    }
    
    print(paste("i", i))
  }
}

names(Parks_grtr_1000)<-gsub("value", "v", names(Parks_grtr_1000))
names(Parks_grtr_1000)<-gsub("lives_saved", "ls", names(Parks_grtr_1000))
names(Parks_grtr_1000)<-gsub("Temp_Day", "bt", names(Parks_grtr_1000))

# Save to shapefile
writeVector(Parks_grtr_1000, "Parks_grtr_1000_with_values_25042024.shp")
#st_write(Parks_grtr_1000, "Parks_grtr_1000_with_value.shp")
```

## Present deaths prevent descriptives
```{r results_descriptives, echo = FALSE}

lives_saved_cols<-grep("ls_", names(Parks_grtr_1000))

# Calculate the total lives saved for each greenspace for all hot days
total_lives_saved<-rowSums(as.data.frame(Parks_grtr_1000[,lives_saved_cols]), na.rm = TRUE)

rm(lives_saved_cols)

Parks_grtr_1000$ttl_ls<-total_lives_saved
rm(total_lives_saved)
 
print("Total lives saved")
sum(total_lives_saved)
print("Maximum lives saved per green space")
max(total_lives_saved)

# Calculate the deaths prevented for each greenspace as number of deaths prevented per m2 of greenspace area
Parks_grtr_1000$ls_p_area<-Parks_grtr_1000$ttl_ls/Parks_grtr_1000$area_calc
 
value_cols<-grep("v_", names(Parks_grtr_1000))
# Calculate the total value for each green space for all hot days
Parks_grtr_1000$ttl_val<-rowSums(as.data.frame(Parks_grtr_1000[,value_cols]), na.rm = TRUE)
rm(value_cols)

print("Total value")
sum(Parks_grtr_1000$ttl_val)

print("Max value per green space")
max(Parks_grtr_1000$ttl_val)

# Calculate the value per unit area
Parks_grtr_1000$v_p_area<-Parks_grtr_1000$ttl_val/Parks_grtr_1000$area_calc

Parks_grtr_1000$area_km2<-Parks_grtr_1000$area_calc/1000000
Parks_grtr_1000$ls_p_km2<-Parks_grtr_1000$ttl_ls/Parks_grtr_1000$area_km2
Parks_grtr_1000$v_p_km2<-Parks_grtr_1000$ttl_val/Parks_grtr_1000$area_km2

Parks_grtr_1000$area_cat<-as.factor(Parks_grtr_1000$area_cat) %>%
  factor(levels = c("1,000 - <5,000", "5,000 - <10,000", "10,000 - <100,000",   "100,000 - <1,000,000", ">=1,000,000"  ))

# Prepare descriptives table
descrptives<-Parks_grtr_1000 %>%
  as.data.frame() %>%
   group_by(area_cat) %>%
   summarise(ls_median=median(ttl_ls, na.rm = TRUE),ls_mean=mean(ttl_ls, na.rm = TRUE), ls_km_median=median(ls_p_km2, na.rm = TRUE),ls_km_mean=mean(ls_p_km2, na.rm = TRUE),v_median=median(ttl_val, na.rm = TRUE),v_mean=mean(ttl_val, na.rm = TRUE), v_km_median=median(v_p_km2, na.rm = TRUE),v_km_mean=mean(v_p_km2, na.rm = TRUE) ) %>%
  as_flextable() %>%
  colformat_double(digits = 4)

# print table
descrptives
rm(descrptives)
#write.csv(descrptives, "lives_saved_val_descriptives.csv")

```

## Plot results
```{r plot_map, echo = FALSE}

# Plot histogram of value
brks<-c(0,  100, 1000, 10000, 100000, 1000000, max(Parks_grtr_1000$ttl_val))

labels<-c("0 - <100", "100 - <1,000", "1,000 - <10,000", "10,000 - <100,000", "100,000 - <1,000,000",">1,000,000")
h<-hist(Parks_grtr_1000$ttl_val, breaks = brks, right = FALSE) 

h_df<-data.frame("Counts"=h$counts, labels)
h_df$labels<-as.factor(h_df$labels)
h_df$labels<-factor(h_df$labels, levels = levels(h_df$labels)[c(2, 5, 3, 4, 6, 1)])

p<-ggplot(h_df, aes(x = labels, y = Counts))+
  geom_col()+
  theme_minimal()+
  xlab("Green space value (€)")+
  ylab("Number of spaces")+
  ylim(c(0, max(pretty(h_df$Counts))))+
   theme(axis.text.x = element_text(angle = 15, vjust = 1, hjust=0.5))
p
setwd(figure_d)
ggsave("hist_value.png", width = 7, height = 3.5, units ="in")

Parks_grtr_1000$Value_cats<-cut(Parks_grtr_1000$ttl_val,breaks=brks,labels = labels, right = FALSE,
                           include.lowest =TRUE)

rm(p, h_df, h, brks, labels)

colour_pal<-brewer.pal(8, "RdYlBu")
setwd(data_d)
central_depts<-vect("central_departments.shp")
crs(central_depts)<-crs(Parks_grtr_1000)

# Plot map of green spaces and their value
p<-ggplot()+
   geom_spatvector(data = Parks_grtr_1000,  aes(fill = Value_cats, colour = Value_cats))+
  geom_spatvector(data = central_depts, colour = "black", fill = NA)+
   theme_minimal()+
  scale_fill_manual(name = "Value (€)",  values = colour_pal[c(1:3, 6:8)])+
  scale_colour_manual(name = "Value (€)",  values = colour_pal[c(1:3, 6:8)])

p
setwd(figure_d)
ggsave("map_values.png", width = 7, height = 3.5, units ="in", dpi = 300)
```