---
title: "Step_1_modelling"
author: "Jo Garrett"
date: '2023-12-21'
output: word_document
---
```{r libraries, message=FALSE, echo=FALSE, warning=FALSE}
## Make sure all libraries are installed and set working directory locations to own locations
library(dplyr)
library(plyr)
library(huxtable)
library(ggplot2)
library(tidyr)
library(gtsummary)
library(tibble)
library(rsq)
library(lmerTest)
library(mgcv)
library(cowplot)
library(lmtest)
library(gratia)
library(terra)
library(tidyterra)

## set data working directory to own location
data_d<-"E:\\Projects\\Users\\JoGarrett\\REGREEN_crib\\Analysis_2024"
## set working directory for output figures to own location
figure_d<-"E:\\Projects\\Users\\JoGarrett\\REGREEN_crib\\Analysis_2024\\Figs_for_paper"
```
## Step 1 - modelling


```{r table_functions, echo = FALSE}

## Create custom functions for presenting model results
glm_tbl_regression<-function(model_output){
  tbl_regression(model_output,
                 intercept = TRUE,
                 estimate_fun= function(x)
                   style_number(x, digits = 2), pvalue_fun=function(x) style_number(x, digits = 3)) %>%
    bold_p() %>%
    add_significance_stars(hide_ci = FALSE, hide_se= TRUE) %>%
    add_glance_table(label = list(AIC~"AIC", Rsq~"R2"),
                     include=c(AIC, nobs, Rsq),
                     glance_fun = function(x) glance(x) %>%
                       add_column(Rsq= rsq(x)))
}

my_tidy <- function(model_output) {
tidy<-broom.mixed::tidy(model_output, conf.int = TRUE,  exponentiate = FALSE, conf.level = 0.95) %>%
  select(-effect, -group, -df) %>%
  filter(!row_number() %in% c(4:5))
return(tidy)

}
lmer_tbl_regression<-function(model_output){
  tbl_regression(model_output,
                 tidy_fun = broom.mixed::tidy,
                 intercept = TRUE,
                 estimate_fun= function(x)
                   style_number(x, digits = 2)) %>%
   add_significance_stars(hide_ci = FALSE, hide_se= TRUE) %>%
    add_glance_table(label = list(AIC~"AIC", Rsq~"R2"),
                     include=c(AIC, nobs, Rsq),
                     glance_fun = function(x) glance(x) %>%
                       add_column(Rsq= unlist(rsq(x)[1])))
}

```


```{r read_data, echo = FALSE}

## Read in data
setwd(data_d)
step_1_results<-read.csv("step_1_resuts_575_150m_15042024.csv")
```

## Descriptives

```{r descriptives_area_ndvi, echo = FALSE}
print("Range green space area")
range(step_1_results$park_area)
print("Range green space NDVI")
range(step_1_results$NDVI)

## Create categories of greenspace area
brks<-c( 1000,5000, 10000, 100000,
        max(step_1_results$park_area, na.rm = TRUE))
labels<-c( "1,000 - <5,000",
          "5,000 - <10,000",
          "10,000 - <100,000","100,000 - <1,000,000")

step_1_results$area_cat<-cut(step_1_results$park_area,breaks=brks,labels = labels, right = FALSE,
                           include.lowest =TRUE)

rm(brks, labels)

## Create categories for NDVI value
brks<-c( 0.2, 0.3,0.4,0.5,
        max(step_1_results$NDVI, na.rm = TRUE))
labels<-c( "0.2 - <0.3", "0.3 - <0.4", "0.4 - <0.5", "0.5 - <0.65")

step_1_results$NDVI_cat<-cut(step_1_results$NDVI,breaks=brks,labels = labels, right = FALSE,
                           include.lowest =TRUE)

rm(brks, labels)

## calculate descriptives two-way table for area and NDVI
tbl_area_ndvi<-table(step_1_results$area_cat, step_1_results$NDVI_cat) %>%
  as.data.frame() %>%
  pivot_wider(names_from = "Var2", values_from = "Freq") %>%
  as_huxtable()

tbl_area_ndvi

rm(tbl_area_ndvi)
```
### Which spaces have no cooling?
```{r spaces_no_cooling, echo = FALSE}

## find which spaces have exhibit no cooling effect 
ind_0_191<-which(step_1_results$Max_dist_191==0)
ind_0_203<-which(step_1_results$Max_dist_203==0)
ind_0_205<-which(step_1_results$Max_dist_205==0)

print("Number of spaces with no cooling on 10th July")
length(ind_0_191)

print("Percentage of spaces with no cooling on 10th July")
length(ind_0_191)/NROW(step_1_results)*100
print("Number of spaces with no cooling on 22nd July")
length(ind_0_203)
print("Percentage of spaces with no cooling on 22nd July")
length(ind_0_203)/NROW(step_1_results)*100
print("Number of spaces with no cooling on 24th July")
length(ind_0_205)
print("Percentage of spaces with no cooling on 24th July")
length(ind_0_205)/NROW(step_1_results)*100

rm(ind_0_191, ind_0_203, ind_0_205)

```

### NDVI descriptive plot
```{r descriptives, echo = FALSE, eval = FALSE}

## Present boxplot of NDVI values by greenspace area category
mean_NDVI <- ddply(step_1_results, "area_cat", summarise, Mean.NDVI=mean(NDVI), sd.NDVI = sd(NDVI), median.ndvi = median(NDVI), lwr.prc=quantile(NDVI, probs = 0.25),upr.prc=quantile(NDVI, probs = 0.75) )
mean_NDVI
rm(mean_NDVI)

p<-ggplot(data=step_1_results, aes(x =area_cat, y = NDVI ))+
  geom_boxplot(notch = FALSE)+
  theme_minimal() +
  ylab("NDVI")+
  xlab(expression("Greenspace area (m"^2~")"))+
  theme(axis.text.x = element_text(angle = 20, vjust =0.5, hjust = 0.5))
p
setwd(figure_d)
ggsave("boxplot_ndvi.png", dpi = 600, width = 4.5, height = 3.5)
rm( p)
```



###  Cooling distance and temperature difference descriptives

```{r descriptives_cooling, echo = FALSE}
print("Median maximum cooling distance 10th July")
median(step_1_results$Max_dist_191)
print("Median maximum cooling distance 22nd July")
median(step_1_results$Max_dist_203)
print("Median maximum cooling distance 24th July")
median(step_1_results$Max_dist_205)

print("Maximum maximum cooling distance 10th July")
max(step_1_results$Max_dist_191)
print("Maximum maximum cooling distance 22nd July")
max(step_1_results$Max_dist_203)
print("Maximum maximum cooling distance 24th July")
max(step_1_results$Max_dist_205)

print("Median temperature difference 10th July")
median(step_1_results$temp_diff_191)
print("Median temperature difference 22nd July")
median(step_1_results$temp_diff_203)
print("Median temperature difference 24th July")
median(step_1_results$temp_diff_205)

print("Maximum temperature difference  10th July")
max(step_1_results$temp_diff_191)
print("Maximum temperature difference  22nd July")
max(step_1_results$temp_diff_203)
print("Maximum temperature difference  24th July")
max(step_1_results$temp_diff_205)

## Presents descriptives table for maximum cooling distance
step_1_results_long<-pivot_longer(step_1_results, 
                                  cols = c(Max_dist_191, Max_dist_203, Max_dist_205), names_to = "Day", values_to = "Max_dist")
step_1_results_long$Day<-gsub("Max_dist_", "", step_1_results_long$Day)

cooling_dist_descriptives<-step_1_results_long %>%
  ddply( c("Day","area_cat"), summarise,  grp.median=median(Max_dist)) %>% pivot_wider(names_from = Day, values_from = grp.median) %>%
  as_huxtable()
print("Maximum cooling distance descriptives")
cooling_dist_descriptives

## Presents descriptives table for temperature difference
step_1_results_long<-pivot_longer(step_1_results, 
                                  cols = c(temp_diff_191, temp_diff_203, temp_diff_205), names_to = "Day", values_to = "Temp_diff")
step_1_results_long$Day<-gsub("temp_diff_", "", step_1_results_long$Day)

temp_diff_descriptives<-step_1_results_long %>%
  ddply( c("Day","area_cat"), summarise,  grp.median=median( Temp_diff)) %>%
 pivot_wider(names_from = Day, values_from = c(grp.median)) %>%
  as_huxtable()
number_format(temp_diff_descriptives)<-3

print("Temperature difference descriptives")
temp_diff_descriptives

rm(temp_diff_descriptives, cooling_dist_descriptives, step_1_results_long)
```

### Distribution of modelling variables
```{r modelling_distance, echo = FALSE}

plot(density(step_1_results$Max_dist_191), main = "Distribution of maximum cooling distance")
plot(density((log10(step_1_results$Max_dist_191)+1)))

plot(density(step_1_results$park_area), main = "Distribution of park area")
plot(density(log10(step_1_results$park_area)), main ="log10 transformed distribution of park area")

plot(density(step_1_results$NDVI))
```


## Modelling
### Generalised additive model
```{r gam_relationship, echo = FALSE}

## Runs GAM predicting maximum cooling distance (transformed) for one day by green space area and NDVI
gam_model<-gam(log10(Max_dist_191+1)~ s(log10(park_area)) + s(NDVI), data = step_1_results)

## Presents plot predicting outcome by area
p1 <- gratia::draw(gam_model, select = "s(log10(park_area))") +
   theme_minimal()+
   xlab(expression("log"[10]*"(Area (m"^2*"))"))+
   ggtitle("a")
 
## Presents plot predicting outcome by NDVI
p2 <- gratia::draw(gam_model, select = "s(NDVI)") +
   theme_minimal()+
   xlab("NDVI")+
   ggtitle("b")
 
plot_grid(p1, p2)
setwd(figured_d)
ggsave("gam_plots_maximum_cool_dist.png", width = 7.5, height = 4, units ="in")
summary(gam_model)

rm(gam_model, p1, p2)

print("Gam model results indicate a quadratic relationship for NDVI only")
```

### Linear modelling- maximum cooling distance

```{r linear_model, echo = FALSE}
## Prepares data for modelling
step_1_results_long_dist<-pivot_longer(step_1_results,
                                  cols = c(Max_dist_191, Max_dist_203,Max_dist_205),
                                  names_to = "Day", values_to ="Max_dist",  names_prefix = "Max_dist_")

## Models maximum cooling distance by green space area and NDVI including a quadratic relationship for NDVI and interactions between both predictors and day 
lm_cd_int<- lm(log10(Max_dist+1) ~ log10(park_area)+ log10(park_area):Day + NDVI + I(NDVI^2)+ Day + I(NDVI^2):Day + NDVI:Day,data = step_1_results_long_dist)

## Prepare results table
lm_cd_int_tbl <- glm_tbl_regression(lm_cd_int)

## Models maximum cooling distance by green space area and NDVI including a quadratic relationship for NDVI for each day separately
lm_cd_191<-lm(log10(Max_dist_191+1)~(log10(park_area)) + NDVI + I(NDVI^2),data=step_1_results)
lm_cd_203<-lm(log10(Max_dist_203+1)~(log10(park_area))  + NDVI + I(NDVI^2),data=step_1_results)
lm_cd_205<-lm(log10(Max_dist_205+1)~(log10(park_area))  + NDVI + I(NDVI^2),data=step_1_results)

## Prepare results tables
lm_cd_191_tbl<-glm_tbl_regression(lm_cd_191)
lm_cd_203_tbl<-glm_tbl_regression(lm_cd_203)
lm_cd_205_tbl<-glm_tbl_regression(lm_cd_205)

## Models maximum cooling distance by green space area and NDVI including a quadratic relationship for NDVI with a random effect for day
lm_cd_re<-lmer(log10(Max_dist+1)~(log10(park_area)) +   NDVI + I(NDVI^2) + (1| Day),data=step_1_results_long_dist)

## Prepare results table
lm_cd_re_tbl<-lmer_tbl_regression(lm_cd_re)

## merge all maximum cooling distance model results tables together
t_merge<-tbl_merge(list(lm_cd_int_tbl,lm_cd_re_tbl, lm_cd_191_tbl, lm_cd_203_tbl, lm_cd_205_tbl), tab_spanner=c("With int","With RE", "Day_191", "Day_203", "Day_205")) %>%
  modify_table_body(~.x %>% arrange(row_type=="glance_statistic")) %>%
  as_flex_table() 

# print maximum cooling distance model results table
t_merge
rm(t_merge, lm_cd_191_tbl, lm_cd_203_tbl, lm_cd_205_tbl,  lm_cd_int_tbl)

# Print equation predicting maximum cooling distance by predictors and their coefficients
print(paste("log10(Maximum cooling distance (m) +1) = (", coef(lm_cd_re)[[1]][1,2]," x log10(area (m^2))) + (", (coef(lm_cd_re)[[1]][1,3]), " X NDVI) + (", coef(lm_cd_re)[[1]][1,4], " X (NDVI)^2) ",coef(lm_cd_re)[[1]][1,1], sep = ""))

rm(lm_cd_191, lm_cd_203, lm_cd_205, lm_cd_int)

```

### Linear modelling-  temperature difference
```{r modelling_temp_diff, echo = FALSE}

## Prepare data for modelling
step_1_results_long_temp<-pivot_longer(step_1_results,cols = c(temp_diff_191, temp_diff_203, temp_diff_205), names_to = "Day", values_to ="temp_diff",  names_prefix = "temp_diff_")

# Check distributions
plot(density(step_1_results_long_temp$temp_diff))
plot(density(log10(step_1_results_long_temp$temp_diff)))
print("transforming by log10 does not really improve the distribution, therefore, will retain as untransformed")

## Exploring nature of relationship between green space area and NDVI and the outcome temperature difference
gam_model<-gam(temp_diff_191~ s(log10(park_area)) + s(NDVI), data = step_1_results)

p1 <- gratia::draw(gam_model, select = "s(log10(park_area))") +
   theme_minimal()+
   xlab(expression("log"[10]*"(Area (m"^2*"))"))+
   ggtitle("c")
 
p2 <- gratia::draw(gam_model, select = "s(NDVI)") +
   theme_minimal()+
   xlab("NDVI")+
   ggtitle("d")
 
plot_grid(p1, p2)
setwd(figure_d)
ggsave("gam_plots_temp_diff.png", width = 7.5, height = 4, units ="in")

summary(gam_model)
rm(gam_model, p1, p2)
print("Gam results indicate a quadratic relationship for both NDVI and park area")

print("Let's test whether including quadratics fit better")
lm_td_191_none<-lm(temp_diff_191~(log10(park_area))  + NDVI, data=step_1_results)
lm_td_191_park<-lm(temp_diff_191~(log10(park_area)) + I((log10(park_area))^2) + NDVI, data=step_1_results)
lm_td_191_ndvi<-lm(temp_diff_191~(log10(park_area)) + NDVI + I(NDVI^2), data=step_1_results)
lm_td_191_all<-lm(temp_diff_191~(log10(park_area)) + I((log10(park_area))^2) + NDVI + I(NDVI^2), data=step_1_results)
lm_td_191_quad_park_only<-lm(temp_diff_191~  I((log10(park_area))^2) + NDVI + I(NDVI^2), data=step_1_results)
AIC_models<-c(AIC(lm_td_191_none), AIC(lm_td_191_park), AIC(lm_td_191_ndvi), AIC(lm_td_191_all))
AIC_models
print("With both fits best and is significiantly different")
rm(AIC_models)
lrtest(lm_td_191_none, lm_td_191_park, lm_td_191_ndvi, lm_td_191_all)
rm(lm_td_191_none, lm_td_191_park, lm_td_191_ndvi, lm_td_191_all, lm_td_191_quad_park_only)

print("They are all significantly different and the one with both NDVI and park area having quadratics fits better")

## Linear model predicting temperature difference between greenspace area and NDVI. A quadratic term is included for both green space area and NDVI. An interaction term is included for day
lm_td_int<- lm(temp_diff~ log10(park_area)+ log10(park_area):Day + I((log10(park_area))^2) + I((log10(park_area))^2):Day + NDVI + I(NDVI^2) +  NDVI:Day + Day + I(NDVI^2):Day,data = step_1_results_long_temp)
## prepare table 
lm_td_int_tbl <- glm_tbl_regression(lm_td_int)

## Linear model predicting temperature difference between greenspace area and NDVI. A quadratic term is included for both green space area and NDVI for each day separately
lm_td_191<-lm(temp_diff_191~(log10(park_area)) + I((log10(park_area))^2) +  NDVI + I(NDVI^2), data=step_1_results)
lm_td_203<-lm(temp_diff_203~(log10(park_area)) + I((log10(park_area))^2) +  NDVI + I(NDVI^2),data=step_1_results)
lm_td_205<-lm(temp_diff_205~(log10(park_area)) + I((log10(park_area))^2) +  NDVI + I(NDVI^2),data=step_1_results)
## Prepare model results tables
lm_td_191_tbl<-glm_tbl_regression(lm_td_191)
lm_td_203_tbl<-glm_tbl_regression(lm_td_203)
lm_td_205_tbl<-glm_tbl_regression(lm_td_205)

## Linear model predicting temperature difference by greenspace area and NDVI includingquadratic terms for both greenspace and NDVI and a random effect for day. 
lm_td_re<-lmer(temp_diff~(log10(park_area)) +I((log10(park_area))^2) +  NDVI + I(NDVI^2) + (1| Day),data=step_1_results_long_temp)
# Prepare model results table
lm_td_re_tbl<-lmer_tbl_regression(lm_td_re)

## Combine model results tables predicting temperature difference
t_merge<-tbl_merge(list(lm_td_int_tbl,lm_td_re_tbl, lm_td_191_tbl, lm_td_203_tbl, lm_td_205_tbl), tab_spanner=c("With int","With RE", "Day_191", "Day_203", "Day_205")) %>%
  modify_table_body(~.x %>% arrange(row_type=="glance_statistic")) %>%
  as_flex_table() 
# print full model results table
t_merge

print("Table for manuscript")
t_merge<-tbl_merge(list(lm_cd_re_tbl,lm_td_re_tbl), tab_spanner=c("Max cooling dist", "Temp diff")) %>%
  modify_table_body(~.x %>% arrange(row_type=="glance_statistic")) %>%
  as_flex_table() 

t_merge
rm(t_merge, lm_td_191_tbl, lm_td_203_tbl, lm_td_205_tbl, lm_td_int_tbl, lm_td_re_tbl,  lm_cd_re_tbl)

# print equation predicting temperature difference by greenspace area and NDVI including coefficients
print(paste0("Temperature difference (°C) = (", coef(lm_td_re)[[1]][1,2], " x log10(area (m^2)) + (", coef(lm_td_re)[[1]][1,3]," x log10(area^2 (m^2))) + (",   coef(lm_td_re)[[1]][1,4], " x NDVI) + (", coef(lm_td_re)[[1]][1,5], " x NDVI^2) + ",  coef(lm_td_re)[[1]][1,1])) 

rm(lm_td_191, lm_td_203, lm_td_205, lm_td_int)
```

## Plots
```{r plotting, echo = FALSE}
color_pal<-c("brown4", "brown2", "darkgoldenrod1", "dodgerblue", "darkblue")

p_all_dist_by_area<-ggplot(step_1_results_long_dist, aes(x = park_area, y = Max_dist, colour = Day))+
 geom_point(aes(shape = Day),  size = 1)+
  theme_minimal()+
  scale_colour_manual(values = rev(color_pal[c(2,4,5)]), labels = c("10th July", "22nd July", "24th July"))+
   scale_shape_manual(values=c(1, 3,4), labels = c("10th July", "22nd July", "24th July"))+
  xlab(bquote('Area '(m^2)))+
  ylab("Maximum cooling distance (m)")+
   scale_x_log10(limits = c(800,  600000), labels = scales::label_comma()) +
    ylim(c(0, max(pretty(step_1_results_long_dist$Max_dist))))+
  ggtitle("a)")


p_all_dist_by_NDVI<-ggplot(step_1_results_long_dist, aes(x = NDVI, y = Max_dist, colour = Day))+
 geom_point(aes(shape = Day),  size = 1)+
  theme_minimal()+
  scale_colour_manual(values = rev(color_pal[c(2,4,5)]), labels = c("10th July", "22nd July", "24th July"))+
   scale_shape_manual(values=c(1, 3,4), labels = c("10th July", "22nd July", "24th July"))+
  xlab("NDVI")+
  ylab("Maximum cooling distance (m)")+
  xlim(c(min(pretty(step_1_results_long_dist$NDVI)), max(pretty(step_1_results_long_dist$NDVI))))+
  ylim(c(0, max(pretty(step_1_results_long_dist$Max_dist)))) +
  ggtitle("b)")


p_all_temp_by_area<-ggplot(step_1_results_long_temp, aes(x = park_area, y = temp_diff, colour = Day))+
 geom_point(aes(shape = Day),  size = 1)+
  theme_minimal()+
  scale_colour_manual(values = rev(color_pal[c(2,4,5)]), labels = c("10th July", "22nd July", "24th July"))+
   scale_shape_manual(values=c(1, 3,4), labels = c("10th July", "22nd July", "24th July"))+
    xlab(bquote('Area '(m^2)))+
  ylab("Temperature difference (°C)")+
   scale_x_log10(limits = c(800,  600000), labels = scales::label_comma()) +
    ylim(c(0, max(pretty(step_1_results_long_temp$temp_diff)))) +
  theme(plot.margin = unit(c(0.5,0.5,0.5,1), "cm"))+
  ggtitle("c)")


p_all_temp_by_NDVI<-ggplot(step_1_results_long_temp, aes(x = NDVI, y = temp_diff, colour = Day))+
 geom_point(aes(shape = Day),  size = 1 )+
 # geom_point(position = "jitter",)+
  theme_minimal()+
  scale_colour_manual(values = rev(color_pal[c(2,4,5)]), labels = c("10th July", "22nd July", "24th July"))+
   scale_shape_manual(values=c(1, 3,4), labels = c("10th July", "22nd July", "24th July"))+
  xlab("NDVI")+
  ylab("Temperature difference (°C)")+
  xlim(c(min(pretty(step_1_results_long_temp$NDVI)), max(pretty(step_1_results_long_temp$NDVI))))+
  ylim(c(-0.05, max(pretty(step_1_results_long_temp$temp_diff))))+
  theme(plot.margin = unit(c(0.5,0.5,0.5,1), "cm")) +
  ggtitle("d)")

#structure of equation for line
#line_dist_area<-(coef_cd_area*(log10(all_days$Area))) + cd_intercept + (coef_cd_ndvi*0.35)
# Model coefficients predicting cooling distance
coef_cd_area<-coef(lm_cd_re)[[1]][1,2]
coef_cd_ndvi<-coef(lm_cd_re)[[1]][1,3]
coef_cd_ndvi2<-coef(lm_cd_re)[[1]][1,4]
cd_intercept<-coef(lm_cd_re)[[1]][1,1]

# Model coefficients predicting temperature difference
coef_td_area<-coef(lm_td_re)[[1]][1,2]
coef_td_area2<-coef(lm_td_re)[[1]][1,3]
coef_td_ndvi<-coef(lm_td_re)[[1]][1,4]
coef_td_ndvi2<-coef(lm_td_re)[[1]][1,5]
td_intercept<-coef(lm_td_re)[[1]][1,1]


# Generating line predicting cooling distance by area
line_dist_area<- (coef_cd_area*(log10(step_1_results_long_dist$park_area))) + (coef_cd_ndvi*0.4)+ (coef_cd_ndvi2*((0.4)^2)) + cd_intercept
line_dist_area<-10^line_dist_area
line_dist_area<-line_dist_area-1

# Generating line predicting cooling distance by NDVI
line_dist_ndvi<- (coef_cd_area*(log10(4000))) + (coef_cd_ndvi*step_1_results_long_dist$NDVI)+ (coef_cd_ndvi2*((step_1_results_long_dist$NDVI)^2)) + cd_intercept
line_dist_ndvi<-10^line_dist_ndvi
line_dist_ndvi<-line_dist_ndvi-1

# Generating line predicting temperature difference by area
line_temp_area<-(coef_td_area*(log10(step_1_results_long_dist$park_area))) + (coef_td_area2*((log10(step_1_results_long_dist$park_area))^2)) + (coef_td_ndvi*0.4)+ (coef_td_ndvi2*((0.4)^2)) + td_intercept

# Generating line predicting temperature difference by NDVI
line_temp_ndvi<-(coef_td_area*(log10(4000))) + (coef_td_area2*((log10(4000))^2)) + (coef_td_ndvi*step_1_results_long_dist$NDVI)+ (coef_td_ndvi2*((step_1_results_long_dist$NDVI)^2)) + td_intercept

## Add lines to plots
p_all_dist_by_area_line<-p_all_dist_by_area+
  geom_line(aes(x = step_1_results_long_dist$park_area, y =line_dist_area), inherit.aes = FALSE)

p_all_dist_by_ndvi_line<-p_all_dist_by_NDVI+
  geom_line(aes(x = step_1_results_long_dist$NDVI, y =line_dist_ndvi), inherit.aes = FALSE)

p_all_temp_by_area_line<-p_all_temp_by_area+
  geom_line(aes(x = step_1_results_long_dist$park_area, y =line_temp_area), inherit.aes = FALSE)

p_all_temp_by_ndvi_line<-p_all_temp_by_NDVI +
  geom_line(aes(x = step_1_results_long_dist$NDVI, y =line_temp_ndvi), inherit.aes = FALSE)

## Align plots
aligned <- align_plots( p_all_dist_by_area_line, p_all_dist_by_ndvi_line, p_all_temp_by_area_line, p_all_temp_by_ndvi_line ,align = "hv")

## Draw plots
plot_grid(ggdraw(aligned[[1]]) , ggdraw(aligned[[2]]), ggdraw(aligned[[3]]), ggdraw(aligned[[4]])  )
setwd(figure_d)
ggsave("temp_diff_dist_all_days.png",
       dpi = 600, height = 6, 
       width = 9)

rm(aligned, p_all_dist_by_area, p_all_dist_by_area_line, p_all_dist_by_NDVI, p_all_dist_by_ndvi_line, p_all_temp_by_area, p_all_temp_by_area_line, p_all_temp_by_NDVI, p_all_temp_by_ndvi_line, line_dist_area, line_dist_ndvi, line_temp_area, line_temp_ndvi, lm_cd_re, lm_td_re)

###Plot histogram
## Distance
brks<-c(seq(0, max(pretty(step_1_results_long_dist$Max_dist)), 100))

h_191<-hist(step_1_results$Max_dist_191,
        breaks = brks, right = FALSE, plot = FALSE)
h_191<-data.frame("Counts" =h_191$counts, "Lowest_val"=h_191$breaks[1:12], "Day"=rep(191, 12))
h_203<-hist(step_1_results$Max_dist_203,
        breaks = brks, right = FALSE, plot = FALSE)
h_203<-data.frame("Counts" =h_203$counts, "Lowest_val"=h_203$breaks[1:12], "Day"=rep(203, 12))
h_205<-hist(step_1_results$Max_dist_205,
        breaks = brks, right = FALSE, plot = FALSE)
h_205<-data.frame("Counts" =h_205$counts, "Lowest_val"=h_205$breaks[1:12], "Day"=rep(205, 12))
h_dist<-rbind(h_191, h_203, h_205)
h_dist$Day<-as.character(h_dist$Day)

hist_dist<-ggplot(h_dist, aes(x = Lowest_val, y = Counts, color=Day, fill = Day, group = Day)) +
  geom_bar(stat = "identity", position = "dodge")+
  scale_colour_manual(values = rev(color_pal[c(2,4,5)]), labels = c("10th July", "22nd July", "24th July"))+
   scale_fill_manual(values = rev(color_pal[c(2,4,5)]), labels = c("10th July", "22nd July", "24th July"))+
  xlab("Max cooling distance (m)")+
   ylab("Number of green spaces")+
  theme_minimal()+
  ylim(c(0, 325))+
  xlim(c(-50, 1200))

rm( h_191, h_203, h_205, brks,  h_dist)

## Temperature difference histogram
brks<-c(seq(0, max(pretty(step_1_results_long_temp$temp_diff)), 0.25))

h_191<-hist(step_1_results$temp_diff_191,
        breaks = brks, right = FALSE, plot = FALSE)
h_191<-data.frame("Counts" =h_191$counts, "Lowest_val"=h_191$breaks[1:8], "Day"=rep(191, 8))
h_203<-hist(step_1_results$temp_diff_203,
        breaks = brks, right = FALSE, plot = FALSE)
h_203<-data.frame("Counts" =h_203$counts, "Lowest_val"=h_203$breaks[1:8], "Day"=rep(203, 8))
h_205<-hist(step_1_results$temp_diff_205,
        breaks = brks, right = FALSE, plot = FALSE)
h_205<-data.frame("Counts" =h_205$counts, "Lowest_val"=h_205$breaks[1:8], "Day"=rep(205, 8))
h_temp<-rbind(h_191, h_203, h_205)
h_temp$Day<-as.character(h_temp$Day)

hist_temp<-ggplot(h_temp, aes(x = Lowest_val, y = Counts, color=Day, fill = Day, group = Day)) +
  geom_bar(stat = "identity", position = "dodge")+
  scale_colour_manual(values = rev(color_pal[c(2,4,5)]), labels = c("10th July", "22nd July", "24th July"))+
   scale_fill_manual(values = rev(color_pal[c(2,4,5)]), labels = c("10th July", "22nd July", "24th July"))+
  xlab("Temperature difference (°C)")+
     ylab("Number of green spaces")+
  theme_minimal()+
  ylim(c(0, 350))

plot_grid(hist_dist, hist_temp)
setwd(figure_d)
ggsave("temp_diff_dist_counts_all_days.png",
       dpi = 600, height = 3, 
       width = 9)

rm(brks, h_191, h_205,h_203, h_temp,  hist_dist, hist_temp, step_1_results_long_dist, step_1_results_long_temp, step_1_results)
```


## Present indicative values for cooling by area and NDVI
```{r cooling_indicative_values, echo = FALSE}
area<-c(1000, 10000, 100000)

dist_by_area<- (coef_cd_area*(log10(area))) + (coef_cd_ndvi*0.4)+ (coef_cd_ndvi2*((0.4)^2)) + cd_intercept
dist_by_area<-10^dist_by_area
dist_by_area<-dist_by_area-1

dist_by_area
NDVI<-c(0.2, 0.4, 0.5, 0.6)
dist_by_ndvi<- (coef_cd_area*(log10(4000))) + (coef_cd_ndvi*NDVI)+ (coef_cd_ndvi2*((NDVI)^2)) + cd_intercept

dist_by_ndvi<-10^dist_by_ndvi
dist_by_ndvi<-dist_by_ndvi-1
temp_by_area<-(coef_td_area*(log10(area))) + (coef_td_area2*((log10(area))^2)) + (coef_td_ndvi*0.4)+ (coef_td_ndvi2*((0.4)^2)) + td_intercept
temp_by_area

temp_by_ndvi<-(coef_td_area*(log10(4000))) + (coef_td_area2*((log10(4000))^2)) + (coef_td_ndvi*NDVI)+ (coef_td_ndvi2*((NDVI)^2)) + td_intercept
df_area<-data.frame(area,dist_by_area, temp_by_area )
df_area<-as_huxtable(df_area)
number_format(df_area)<-3
df_area

df_ndvi<-data.frame(NDVI,dist_by_ndvi, temp_by_ndvi) %>%
  as_huxtable()
number_format(df_ndvi)<-3
df_ndvi
rm(area, dist_by_area, NDVI, dist_by_ndvi, df_area, df_ndvi, temp_by_ndvi, temp_by_area)

```

## Step 2 - upscale to all greenspaces
### Read in parks data

```{r read_parks, echo = FALSE}

## Set working directory
setwd(data_d)

## Load in parks data using terra package
Parks_master<-vect("greenspaces_central.shp")

## set coordinate reference system
crs(Parks_master)<-"EPSG:2154"

## Calculate the area of each green space
Parks_master$area_calc<-expanse(Parks_master)

brks<-c(0, 1, 10,50, 100, 1000,5000, 10000, 100000, 1000000,
        max(Parks_master$area_calc, na.rm = TRUE))
labels<-c("0 - <1", "1 - <10", "10 - <50","50 - <100", 
          "100 - <1,000", "1,000 - <5,000",
          "5,000 - <10,000",
          "10,000 - <100,000","100,000 - <1,000,000",
          ">=1,000,000")

## Label each green space with a category based on area
Parks_master$area_cat<-cut(Parks_master$area_calc,
                           breaks=brks,
                   labels = labels,
                           right = FALSE,
                           include.lowest =TRUE)

rm(brks, labels)

ind<-which(Parks_master$area_calc>1000)
## Retain only green spaces which are at least 1000m2
Parks_grtr_1000<-Parks_master[ind,]
rm(Parks_master,ind)

## Create a rectangle with which to crop the temperature rasters. 
## This is created from a bounding box of the green spaces plus a buffer of 1000 m
parks_bounding<-ext(Parks_grtr_1000) %>%
  as.polygons() %>%
  buffer(width = 1000)
```

### Read NDVI

```{r read_NDVI, echo = FALSE}

## Read in NDVI and attach to parks data
NDVI<-rast("2019191_interpolated_NDVI.tif") %>%
  terra::project( "EPSG:2154") %>%
  crop( parks_bounding) %>%
  disagg(fact= 6) %>%
  ## Replace values that are greater than 0.75 with NA
  {.[. > 0.75] <- NA; .}
rm( parks_bounding)

# Calculate mean NDVI value for each green space
NDVI_mean <- terra::extract(NDVI,
                            Parks_grtr_1000, fun = function(x) mean(x, na.rm = TRUE), ID = FALSE)
Parks_grtr_1000$NDVI<-NDVI_mean[,1]
rm(NDVI, NDVI_mean)
```

###  Calculate maximum cooling distance for all green spaces >1,000m2
```{r calc_max_cooling_dist}

# Calculate maximum cooling distance using the model coefficients, and the greenspace area and NDVI value
line_dist_area<- (coef_cd_area*(log10(Parks_grtr_1000$area_calc))) + (coef_cd_ndvi*Parks_grtr_1000$NDVI)+ (coef_cd_ndvi2*((Parks_grtr_1000$NDVI)^2)) + cd_intercept
line_dist_area<-10^line_dist_area
line_dist_area<-line_dist_area-1

Parks_grtr_1000$pred_cooling_dist<-line_dist_area
rm(line_dist_area)

# Calculate maximum cooling distance using the model coefficients, and the greenspace area and NDVI value
line_temp<-(coef_td_area*(log10(Parks_grtr_1000$area_calc))) + (coef_td_area2*((log10(Parks_grtr_1000$area_calc))^2)) + (coef_td_ndvi*Parks_grtr_1000$NDVI)+ (coef_td_ndvi2*((Parks_grtr_1000$NDVI)^2)) + td_intercept
Parks_grtr_1000$pred_temp_diff<-line_temp
rm(line_temp)

print("Median predicted cooling distance for all green spaces")
median(Parks_grtr_1000$pred_cooling_dist)

print("Maximum predicted cooling distance for all green spaces")
max(Parks_grtr_1000$pred_cooling_dist)

print("Median predicted temperature difference for all green spaces")
median(Parks_grtr_1000$pred_temp_diff)

print("Maximum predicted temperature difference for all green spaces")
max(Parks_grtr_1000$pred_temp_diff)

```


### Present histogram of predicted values
```{r hist_preds, echo = FALSE}
bins_dist<-c(-1,20, 50, 100, 200, 400, 600, 1550)
h<-hist(Parks_grtr_1000$pred_cooling_dist, breaks = bins_dist, right = FALSE, plot = FALSE)

bin_labels<-character(length = length(bins_dist)-1)
#[a, b): Includes a but not b.
for (i in 1:(length(bins_dist)-1)){
  bin_labels[i]<-paste0(bins_dist[i], " - <", bins_dist[i+1])
}
rm(i)

h_df<-data.frame(bin_labels, "Counts"=h$counts)
h_df$bin_labels<-factor(h_df$bin_labels, levels =unique(h_df$bin_labels))

p1<-ggplot(h_df, aes(x = bin_labels, y =Counts ))+
  geom_col()+
  theme_minimal()+
  xlab("Maxmium cooling distance (m)")+
  ylab("Count")+
  theme(text = element_text(size = 17), axis.text.x = element_text(angle = 30, vjust = 1, hjust=0.5))+
  ggtitle("a)")

bins_dist<-c(0,0.1, 0.2,0.3, 0.4, 0.5, 1, 2.7)
h<-hist(Parks_grtr_1000$pred_temp_diff, breaks = bins_dist, right = FALSE, plot = FALSE)

bin_labels<-character(length = length(bins_dist)-1)
#[a, b): Includes a but not b.
for (i in 1:(length(bins_dist)-1)){
  bin_labels[i]<-paste0(bins_dist[i], " - <", bins_dist[i+1])
}
rm(i)

h_df<-data.frame(bin_labels, "Counts"=h$counts)
h_df$bin_labels<-factor(h_df$bin_labels, levels =unique(h_df$bin_labels))

p2<-ggplot(h_df, aes(x = bin_labels, y =Counts ))+
  geom_col()+
  theme_minimal()+
  xlab("Temperature difference (°C)")+
  ylab("Count")+
  theme(text = element_text(size = 17),
        axis.text.x = element_text(angle = 30, vjust = 1, hjust=0.5))+
  ggtitle("b)")

aligned <- align_plots( p1, p2,align = "hv")

plot_grid(ggdraw(aligned[[1]]) , ggdraw(aligned[[2]]) )
setwd(figure_d)
ggsave("pred_cooling_effects_hist.png",
       dpi = 600, height = 4.75, 
       width = 10.5)

rm(aligned, p1, p2, h_df, bin_labels, bins_dist, h)
```

### Descriptives of predictions
```{r descriptive_predictions, echo = FALSE}

#names(Parks_grtr_1000)
Parks_grtr_1000_df<-as.data.frame(Parks_grtr_1000)

summary_pred<-Parks_grtr_1000_df %>%
  dplyr::group_by(area_cat) %>% 
  dplyr::summarise(Median_cd = median(pred_cooling_dist, na.rm = TRUE), Max_cd = max(pred_cooling_dist, na.rm = TRUE),Median_td = median(pred_temp_diff, na.rm = TRUE), Max_td = max(pred_temp_diff, na.rm = TRUE), Count = n()) %>%
  as_huxtable()

number_format(summary_pred)<-3
summary_pred
rm(Parks_grtr_1000_df, summary_pred)

```

## Calculate population benefitting

### Read in population data
```{r population_data, echo = FALSE}
## Population data

setwd(data_d)

Pop<-vect("densibati_central_ile_de_france.shp") %>%
  makeValid()

total_pop<-sum(Pop$POPMEN2016,na.rm = TRUE)

```

```{r parks_max_cd_buff, echo = FALSE}

setwd(data_d)
writeVector(Parks_grtr_1000, "Parks_grtr_1000_with_preds.shp")
#Parks_grtr_1000<-vect("Parks_grtr_1000_with_preds.shp")

## Calculate buffers for all green spaces for a width of the maximum cooling distance
Parks_grtr_1000_with_cd_buffs<-terra::buffer(Parks_grtr_1000, width=Parks_grtr_1000$pred_cooling_dist) %>%
  terra::aggregate() 

writeVector(Parks_grtr_1000_with_cd_buffs,"Parks_grtr_1000_with_cd_buffs.shp" , overwrite = TRUE)

# retain only buildings which intersect with the cooling areas
is_related_pop<-is.related(Pop, Parks_grtr_1000_with_cd_buffs, relation="intersects")

ind<-which(is_related_pop==TRUE)

Pop_related<-Pop[ind,]
rm(ind)

## Calculate the building area
Pop_related$OG_area<-expanse(Pop_related)
# Intersect the buildings with the cooling areas
Pop_inter<-terra::intersect(Pop_related,Parks_grtr_1000_with_cd_buffs)
# Calculate the area of intersected building area
Pop_inter$Inter_area<-expanse(Pop_inter)
# Calculate the percentage of the intersected area of the original building area
Pop_inter$Area_prc<-Pop_inter$Inter_area/Pop_inter$OG_area
ind<-which(Pop_inter$Area_prc>1)
Pop_inter$Area_prc[ind]<-1
# Calculate the population in the intersected area as a proportion of the original building population by the proportion of the intersected area
Pop_inter$Pop_recalc<-Pop_inter$POPMEN2016*Pop_inter$Area_prc
# Calculate the number of people living within cooling areas
MCD_pop<-sum(Pop_inter$Pop_recalc, na.rm = TRUE)

print("Total population benefiting from cooling")
MCD_pop
print("Percentage benefiting from cooling")
MCD_pop/total_pop*100

rm(Pop_inter, Pop_related, is_related_pop, MCD_pop, total_pop)
```

```{r plotting_pop_MCD, echo = FALSE}


p<-ggplot()+
   geom_spatvector(data = Pop,  aes(fill = "gray90", colour = "gray90"))+
 geom_spatvector(data = Parks_grtr_1000_with_cd_buffs, aes(colour="gray34",fill = "gray80"))+
     geom_spatvector(data = Parks_grtr_1000, aes(colour="gray34",fill = "darkseagreen"))+
    scale_fill_identity(name = NULL,guide = 'legend',labels = c("Public green spaces", "Maximum cooling distance", "Buildings")) +
  theme_minimal()+
    scale_colour_manual(values =c("gray90"="gray90","gray34"="gray34",  "darkseagreen"= "darkseagreen"),guide = 'none', labels = c("Public green spaces", "Maximum cooling distance", "Buildings"))
p

setwd(figure_d)
ggsave("map_greenspaces_mcd.png", dpi = 600, width = 6, height = 5)

```

