---
title: "Step_1_subset_and_modelling"
author: "Jo Garrett"
date: '2023-06-28'
output: 
  officedown::rdocx_document:
    reference_docx: ref_doc.docx
    page_margins:
      bottom: 0.2
      top: 0.5
      right: 0.75
      left: 0.5
      header: 0.2
      footer: 0.2
      gutter: 0.2
---

```{r load_packages, echo = FALSE, warning = FALSE, message = FALSE}

## Make sure all libraries are installed and set data and figure directories to own directories
library(cowplot)
library(dplyr)
library(flextable)
library(forcats)

library(ggplot2)
library(gtsummary)
library(huxtable)
library(lmtest)
library(lmerTest)
library(mgcv)
library(plyr)
library(pracma)
library(readr)
library(rsq)
library(sf)
library(terra)
library(tibble)
library(tidyr)
library(zoo)

data_d<-"E:\\Projects\\Users\\JoGarrett\\REGREEN_crib\\Analysis_2024"
figure_d<-"E:\\Projects\\Users\\JoGarrett\\REGREEN_crib\\Analysis_2024\\Figs_for_paper"
```
## 1. Read in greenspace data

- Read in all green spaces for the central four departments of Paris
- Retain only green spaces >1000m2 for further analyses
- Spaces removed which are planned (not yet in existence)
```{r read_parks, echo = FALSE}

## Set working directory
setwd(data_d)

## Load in parks data using terra package
Parks_master<-vect("greenspaces_central.shp")

brks<-c(0, 1, 10,50, 100, 1000,5000, 10000, 100000, 1000000,
        max(Parks_master$area_terra, na.rm = TRUE))
labels<-c("0 - <1", "1 - <10", "10 - <50","50 - <100", 
          "100 - <1,000", "1,000 - <5,000",
          "5,000 - <10,000",
          "10,000 - <100,000","100,000 - <1,000,000",
          ">=1,000,000")

## Label each green space with a category based on area
Parks_master$area_cat<-cut(Parks_master$area_terra,
                           breaks=brks,
                   labels = labels,
                           right = FALSE,
                           include.lowest =TRUE)

rm(brks, labels)

ind<-which(Parks_master$area_terra>1000)
## Retain only green spaces which are at least 1000m2
Parks_grtr_1000<-Parks_master[ind,]
rm(Parks_master,ind)

## Create a rectangle with which to crop the temperature rasters. 
## This is created from a bounding box of the green spaces plus a buffer of 1000 m
parks_bounding<-ext(Parks_grtr_1000) %>%
  as.polygons() %>%
  buffer(width = 1000)

```

## 2. Read in temperature data for three days - 10th, 22nd, 24th July 2019

```{r read_temp_data, echo =FALSE}

## Set working directory
setwd(data_d)

## Read in data
temp_proj_191<-rast("Temp_data\\2019191_TAVG.tif") %>%
  ## project to the same coordinate reference system as the green spaces
  terra::project( "EPSG:2154") %>%
## Crop to the bounding box 
  crop( parks_bounding) %>%
## Resample by a factor of 6 (to every 5 m)
  disagg( fact= 6)

temp_proj_203<-rast("Temp_data\\2019203_TAVG.tif") %>%
  terra::project( "EPSG:2154") %>%
  crop( parks_bounding) %>%
  disagg( fact= 6)

temp_proj_205<-rast("Temp_data\\2019205_TAVG.tif") %>%
  terra::project( "EPSG:2154") %>%
  crop( parks_bounding) %>%
  disagg( fact= 6)
```

## 3. Read in NDVI

```{r read_NDVI, echo = FALSE}
## Set working directory
setwd(data_d)
## Read in NDVI and attach to parks data
NDVI<-rast("2019191_interpolated_NDVI.tif") %>%
  terra::project( "EPSG:2154") %>%
  crop( parks_bounding) %>%
  disagg(fact= 6) %>%
  ## Replace values that are greater than 0.75 with NA, these are values associated with water
  {.[. > 0.75] <- NA; .}

rm( parks_bounding)

# Calculate mean NDVI per green space
NDVI_mean <- terra::extract(NDVI,
                            Parks_grtr_1000, fun = function(x) mean(x, na.rm = TRUE), ID = FALSE)
Parks_grtr_1000$NDVI<-NDVI_mean[,1]
rm(NDVI, NDVI_mean)
```


## 4. Both distance to green and water

```{r distance_to_others, echo = FALSE, fig.width=9, fig.height = 3.5}

## find the columns which are associated with the distances to the nearest green space and the nearest water body
ind<-which(names(Parks_grtr_1000)=="water_dist" | names(Parks_grtr_1000)=="nearest_di")

distances<-as.data.frame(Parks_grtr_1000[,ind])

## Find the distance to either the nearest other green space or water body, whichever is closest
Parks_grtr_1000$green_or_water_dist<-apply(distances, 1, function(x) min(x, na.rm = TRUE))
rm(distances, ind)


## Produce figure of distances

range(Parks_grtr_1000$water_dist)
brks<-c(0,30, 100,250,500,
        max(Parks_grtr_1000$nearest_di, na.rm = TRUE))
labels<-c("0 -<30", "30 - <100", 
          "100 - <250",  "250 - <500", ">500")

## Calculate table of counts associated with each interval
h_green<-hist(Parks_grtr_1000$nearest_di,
        breaks = brks, right = FALSE, plot = FALSE)
h_water<-hist(Parks_grtr_1000$water_dist,
        breaks = brks, right = FALSE, plot = FALSE)
h_both<-hist(Parks_grtr_1000$green_or_water_dist,
        breaks = brks, right = FALSE, plot = FALSE)

distance_counts_green<-data.frame(labels, h_green$counts)
names(distance_counts_green)[2]<-"Counts"
distance_counts_water<-data.frame(labels, h_water$counts)
names(distance_counts_water)[2]<-"Counts"
distance_counts_both<-data.frame(labels, h_both$counts)
names(distance_counts_both)[2]<-"Counts"

rm(h_both, h_water, h_green)

distance_counts_green$Space<-"Green"
distance_counts_water$Space<-"Water"
distance_counts_both$Space<-"Green_or_water"

distance_counts<-rbind(distance_counts_green, distance_counts_water, distance_counts_both)

rm(distance_counts_green, distance_counts_water, distance_counts_both, brks, labels)

distance_counts$Space<-as.factor(distance_counts$Space)
distance_counts$Space<-factor(distance_counts$Space, levels = c("Green", "Water", "Green_or_water"))
distance_counts$Space<-fct_recode(distance_counts$Space, 'Green or water' = "Green_or_water")

distance_counts$labels<-as.factor(distance_counts$labels)
distance_counts$labels<-factor(distance_counts$labels, levels = c("0 -<30", "30 - <100",  "100 - <250", "250 - <500", ">500"))

p<-ggplot(distance_counts, aes(x = labels, y = Counts))+
  geom_col(show.legend = FALSE)+
  theme_minimal()+
  facet_wrap(~Space)+
  coord_flip() +
  ylim(c(0, max(pretty(distance_counts$Counts)+50)))+
  xlab("Distance to nearest (m)")
p

ggsave(file.path(figure_d, "S5_Paris_region_distance_to_others.png"), dpi = 300, width = 7, height = 2.5)


## print table of distance categories and their counts
distance_counts <-pivot_wider(distance_counts,  values_from = Counts, names_from = Space) %>%
  as_huxtable()
distance_counts
rm(p)
#write.csv(distance_counts, 
 #         "distance_counts_all_spaces.csv")
rm(distance_counts)
```


## Creating green spaces analysis dataset
>1000 sq m and >150 m distance from a green space or water body

```{r analysis_data, echo = FALSE}

## Select green spaces which are at least 150 m from another green space or water body
ind<-which(Parks_grtr_1000$nearest_di>150 & Parks_grtr_1000$water_dist>150)
Parks_master_150m_dist<-Parks_grtr_1000[ind,]

writeVector(Parks_master_150m_dist, "Parks_master_150m_dist.shp", overwrite = TRUE)

range(Parks_master_150m_dist$area_terra)

```


```{r analysis, echo = FALSE}

## Create empty data.frames
Min_dist<-data.frame(Min_dist_191= numeric(NROW(Parks_master_150m_dist)), Min_dist_203= numeric(NROW(Parks_master_150m_dist)), Min_dist_205= numeric(NROW(Parks_master_150m_dist)))

Max_dist<-data.frame(Max_dist_191= numeric(NROW(Parks_master_150m_dist)), Max_dist_203= numeric(NROW(Parks_master_150m_dist)), Max_dist_205= numeric(NROW(Parks_master_150m_dist)))

mean_temp_at_peak<-data.frame(Peak_temp_191= numeric(NROW(Parks_master_150m_dist)), Peak_temp_203= numeric(NROW(Parks_master_150m_dist)), Peak_temp_205= numeric(NROW(Parks_master_150m_dist)))

boundary_temp<-data.frame(boundary_temp_191= numeric(NROW(Parks_master_150m_dist)), boundary_temp_203= numeric(NROW(Parks_master_150m_dist)), boundary_temp_205= numeric(NROW(Parks_master_150m_dist)))

## Set buffer intervals (and create labels) from -15 m fom the greenspace boundary, to 1300 m from the greenspace boundary with intervals of 30 m
brks<-c(-15, seq(0, 1300, 30))
labels<-character(length=(length(brks))-1)
for (i in 1:length(labels)){
  labels[i]<-paste(brks[i], " - <", brks[i+1], sep = "")
}
rm( i) 

## run analysis on each greenspace in the subset for each temperature day
for (i in 1:NROW(Parks_master_150m_dist)){
  mean_temp_191<-numeric(length=(length(brks))-1)
  mean_temp_203<-numeric(length=(length(brks))-1)
  mean_temp_205<-numeric(length=(length(brks))-1)
  ## Run analysis for each 30 m intervals - calculates mean temperature for 30 m buffers around the greenspace
  for (j in 1:44){
    # Create a buffer at the interval minimum
    Buff_min<-terra::buffer(Parks_master_150m_dist[i,], width=brks[j])
    ## Where -15 m from the boundary has zero area, try -10 m and then -5 m to calculate boundary temperature
    if (j == 1 & expanse(Buff_min) == 0){
      Buff_min<-terra::buffer(Parks_master_150m_dist[i,], width=-10)
    }
    if (j == 1 & expanse(Buff_min) == 0){
      Buff_min<-terra::buffer(Parks_master_150m_dist[i,], width=-5)
    }
    # Create a buffer at the interval minimum
    Buff_max<-terra::buffer(Parks_master_150m_dist[i,], width=brks[j+1])
    # Cut out the minimum buffer from the maximum to create a ring 
    Buff_ring<-terra::symdif(Buff_max,Buff_min)
    ## Select only Buffer rings where they are greater than 1 m
    if(NROW(Buff_ring)>1){
      Buff_area<-expanse(Buff_ring)
      Buff_ring<-Buff_ring[Buff_area>1]
    }
    rm(Buff_min, Buff_max)
    ## Calculate mean temperature within the buffer for each of the three hot days
    mean_temp_191[j]<-as.numeric(terra::extract(temp_proj_191, Buff_ring, fun = mean)[2])
    mean_temp_203[j]<-as.numeric(terra::extract(temp_proj_203, Buff_ring, fun = mean)[2])
    mean_temp_205[j]<-as.numeric(terra::extract(temp_proj_205, Buff_ring, fun = mean)[2])
    rm(Buff_ring)
  }
  temp_dist<-data.frame(mean_temp_191,mean_temp_203,mean_temp_205,  brks[2:length(brks)], labels)
  names(temp_dist)[1:4]<-c("Temp_191", "Temp_203", "Temp_205", "Distance")
  rm(mean_temp_191, mean_temp_203, mean_temp_205)
  for (k in 1:3){
    ind<-which(temp_dist[,k]>0)
    # Find maximum cooling distance 
    ## Only for where mean temperature within 0 - 30 m is greater than -15 - 0 m
    if (temp_dist[(min(ind))+1,k]>temp_dist[min(ind),k]){
      ## Calculate location of peaks
      pks<-findpeaks(temp_dist[,k])
      ## Store the minimum and maximum buffer distance of the peak location
      Min_dist[i,k]<-brks[(pks[1,2])]
      Max_dist[i,k]<-brks[(pks[1,2])+1]
      ## Store the peak temperature
      mean_temp_at_peak[i,k]<-pks[1,1]
      rm(pks)
    } else {
      ## If temperature 0 - 30 m is not greater than -15- 0 m, store minimum distance as <0, max distance as 0 and store temperature value
      Min_dist[i,k]<-(-5)
      Max_dist[i,k]<-0
      mean_temp_at_peak[i,k]<-temp_dist[1,k]
    }
  }
  ## Store boundary temperatures
  boundary_temp$boundary_temp_191[i]<-temp_dist$Temp_191[1]
  boundary_temp$boundary_temp_203[i]<-temp_dist$Temp_203[ 1]
  boundary_temp$boundary_temp_205[i]<-temp_dist$Temp_205[1]
  # print(i)
  rm(temp_dist)
}

rm(i,j,k, ind, brks, labels)

## Calculate temperature differences between the peak temperature and the boundary temperature
temp_diff_191<-mean_temp_at_peak$Peak_temp_191-boundary_temp$boundary_temp_191
temp_diff_203<-mean_temp_at_peak$Peak_temp_203-boundary_temp$boundary_temp_203
temp_diff_205<-mean_temp_at_peak$Peak_temp_205-boundary_temp$boundary_temp_205

temp_diff<-data.frame(temp_diff_191, temp_diff_203, temp_diff_205)

rm(temp_diff_191, temp_diff_203, temp_diff_205)


step_1_results<-data.frame(boundary_temp, mean_temp_at_peak, temp_diff, "park_area"=Parks_master_150m_dist$area_terra, Min_dist, Max_dist, "NDVI"=Parks_master_150m_dist$NDVI)

## Save results
write.csv(step_1_results, "step_1_resuts_575_150m_15042024.csv")
rm( boundary_temp, mean_temp_at_peak, temp_diff, Min_dist, Max_dist)

```
